---
title: TITLE HERE
date: "`r Sys.Date()`"
output: 
  html_document:
          code_folding: hide
          code_download: true
          toc: true
          toc_float: true
  github_document:
          html_preview: false
          toc: true
          toc_depth: 3
---

```{r setup, include=T, message = F, warning = F, echo = F}
knitr::opts_chunk$set(
  echo = knitr::is_html_output(),
  fig.align = "center",
  fig.path = "../Figures/markdown/",
  dev = c("png", "pdf"),
  message = FALSE,
  warning = FALSE,
  collapse = T
)

theme_matt = function(base_size = 18,
                      dark_text = "grey20"){
  mid_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[2]
  light_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[3]
  
  ggpubr::theme_pubr(base_family="sans") %+replace% 
    theme(
      panel.background  = element_rect(fill="transparent", colour=NA), 
      plot.background = element_rect(fill="transparent", colour=NA), 
      legend.background = element_rect(fill="transparent", colour=NA),
      legend.key = element_rect(fill="transparent", colour=NA),
      text = element_text(colour = mid_text, lineheight = 1.1),
      title = element_text(size = base_size * 1.5,
                           colour = dark_text),
      axis.text = element_text(size = base_size,
                               colour = mid_text),
      axis.title.x = element_text(size = base_size * 1.2,
                                  margin = unit(c(3, 0, 0, 0), "mm")),
      axis.title.y = element_text(size = base_size * 1.2,
                                  margin = unit(c(0, 5, 0, 0), "mm"), 
                                  angle = 90),
      legend.text = element_text(size=base_size * 0.9),
      legend.title = element_text(size = base_size * 0.9, 
                                  face = "bold"),
      plot.margin = margin(0.25, 0.25, 0.25, 0.25,"cm")
    )
}

theme_matt_facets = function(base_size = 18,
                             dark_text = "grey20"){
  mid_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[2]
  light_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[3]
  
  theme_bw(base_family="sans") %+replace% 
    theme(
      panel.grid = element_blank(),
      panel.background  = element_rect(fill="transparent", colour=NA), 
      plot.background = element_rect(fill="transparent", colour=NA), 
      legend.background = element_rect(fill="transparent", colour=NA),
      legend.key = element_rect(fill="transparent", colour=NA),
      text = element_text(colour = mid_text, lineheight = 1.1),
      strip.text.x = element_text(size = base_size),
      title = element_text(size = base_size * 1.5,
                           colour = dark_text),
      axis.text = element_text(size = base_size,
                               colour = mid_text),
      axis.title.x = element_text(size = base_size * 1.2,
                                  margin = unit(c(3, 0, 0, 0), "mm")),
      axis.title.y = element_text(size = base_size * 1.2,
                                  margin = unit(c(0, 5, 0, 0), "mm"), 
                                  angle = 90),
      legend.text = element_text(size=base_size * 0.9),
      legend.title = element_text(size = base_size * 0.9, 
                                  face = "bold"),
      plot.margin = margin(0.25, 0.25, 0.25, 0.25,"cm")
    )
}
```

## Historical Temperature Data
Daily temperature data was retrieved from the National Centers for Environmental Information from three sites in Boston - Station USW00094701 in Boston, Station USC00191097 in Cambridge, and Station USW00014739 at Boston Logan International Airport. Only maximum and minimum daily temperatures were recorded for the historical data (<1950). To assemble a complete data record, daily maximum and minimum values were first averaged for each day to account for temporal overlap between the data sets. The mid-point between the minimum and maximum was then used a proxy for the daily mean temperature. 

This method produced reasonable estimates for the daily mean temperature. There was a tight correlation between estimated and observed mean daily temperatures. 

```{r}
ggplot(temp_record, aes(x = mean, y = mean_est)) + 
  geom_abline(slope = 1) + 
  geom_point() + 
  geom_smooth() + 
  labs(x = "Observed Mean Temp. (°C)",
       y = "Estimated Mean Temp. (°C)") + 
  theme_matt()
```


```{r}
cor.test(x = temp_record$mean_est, y = temp_record$mean)
```

The temperature record suggests distinct warming in Eastern Massachusetts since the 1880s. 

```{r}
temp_record %>%  
  drop_na(mean_est) %>% 
  group_by(year) %>%  
  summarise(ann_mean = mean(mean_est, na.rm = T),
            ann_max = max(max, na.rm = T),
            ann_min = min(min, na.rm = T)) %>% 
  filter(year != max(year) & year != min(year)) %>% 
  pivot_longer(cols = c(ann_mean:ann_min), 
               names_to = "metric", 
               values_to = "temp") %>% 
  filter(metric == "ann_mean") %>% 
ggplot(aes(x = year, y = temp)) + 
  geom_point() + 
  geom_smooth() +
  labs(x = "Year", 
       y = "Temperature (°C)") + 
  theme_matt() + 
  theme(legend.position = "right")

# ggplot(temp_record, aes(x = date, y = mean_est)) + 
#   geom_line() + 
#   geom_smooth(colour = "red") + 
#   theme_minimal() + 
#   theme(panel.grid = element_blank())
```

```{r}

database %>% 
  select(species, region) %>%
  distinct() %>% 
  group_by(region) %>% 
  count() %>% 
  drop_na() %>% 
  ungroup() %>% 
  mutate(region = fct_reorder(region, n)) %>% 
  ggplot(aes(x = region, y = n)) + 
  geom_bar(stat = 'identity') + 
  theme(axis.text.x = element_text(angle = 300, vjust= 0.5, hjust = 0))

database %>% 
  group_by(species) %>% 
  count() %>% 
  drop_na() %>% 
  ungroup() %>% 
  mutate(species = fct_reorder(species, n)) %>% 
  ggplot(aes(x = species, y = n)) + 
  geom_bar(stat = 'identity') + 
  theme(axis.text.x = element_text(angle = 300, vjust = 0.5, hjust = 0))


database %>% 
  group_by(species, region) %>% 
  count() %>% 
  drop_na() %>% 
  ungroup() %>% 
  count(species) %>% 
  mutate(species = fct_reorder(species, n)) %>% 
  ggplot(aes(x = species, y = n)) + 
  geom_bar(stat = 'identity') + 
  theme(axis.text.x = element_text(angle = 300, vjust = 0.5, hjust = 0))
```

```{r}

# Download County Subdivision shapefiles for MA
# Using county subdivisions (cousub) usually provides better coastal detail
ma_map <- county_subdivisions(state = "MA", cb = TRUE, class = "sf")

ggplot(data = ma_map) +
  geom_sf(fill = "lightblue", color = "lightblue") +
  theme_minimal() +
  geom_point(data = drop_na(database, lat), aes(x = long, y = lat, colour = confidence),
             size = 1) + 
  scale_colour_manual(values = c("high" = "black",
                                 "low" = "grey50",
                                 "medium" = "grey80")) + 
  coord_sf(xlim = c(-69.85, -71.1),
           ylim = c(41.2,42.1)) + 
  labs(x = "Longitude",
       y = "Latitude") +
  theme(panel.grid.major = element_line(color = "gray90", linewidth = 0.5))

```


